---
title: "Lab 3 Part 2"
author: "Yansong Li"
date: "2023-01-28"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = '/tep')
```

## 1. First read in the three csv files: gdppercapitaandgini and airpollution

For the first question, I first set my working directory, then read and stored the files in two variants: AirPollu and GDP. I then changed the titles into shorter ones that are more descriptive and easier to work with. To prepare for following questions, I calculated the outdoor air pollution death rate for China and India and stored the calculated values in a new column called OutDoorPollDeath. I also alculated the Log of Population GDP (Log of base 10).

```{r Q1}
setwd("D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 3/ESP106_week3_lab1")
GDP <- read.csv("gdppercapiandgini.csv")
print(GDP)
AirPollu <- read.csv("airpollution.csv")
print(AirPollu)
colnames(AirPollu) <- c("Country",'Code','Year','PM_Death_Ambient','SolidFuels_Death_HouseHold','Ozone_Death_Ambient','Death_Airpollution')
colnames(GDP) <- c('Country','Code','Year','Total_Population_GDP','Continent','Gini','Outputside_RealGDP_Per_Capita')
AirPollu$OutDoorPolluDeath <- AirPollu$PM_Death_Ambient+AirPollu$Ozone_Death_Ambient
GDP$LogPopGDP <- log10(GDP$Total_Population_GDP)
```

## Chose two countries that you are interested in and make a plot showing the death rates from indoor air pollution and outdoor air pollution (sum of particulate matter and ozone) over time

For this question, I first pulled out the rows of data that are related to China and India from the AirPollu dataframe and stored them into their own respective variables named ChinaAir and IndiaAir. I then I summed the Ambient Ozone Death and the Ambient Particulate matter death of both countries, and assigned them to a new column.To plote a single graph with two countries, I merged the ChinaAir and IndiaAir dataframe by year. After that it was just simply ploting out the graphs and adding the lines.

```{r Q2}
setwd("D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 3/ESP106_week3_lab1")
unique(AirPollu$Entity)
ChinaAir <- AirPollu[AirPollu$Country=="China",]
IndiaAir <- AirPollu[AirPollu$Country=='India',]
ChinaAir$China_outdoor_death <-ChinaAir$Ozone_Death_Ambient+ChinaAir$PM_Death_Ambient
IndiaAir$India_outdoor_death <-IndiaAir$Ozone_Death_Ambient+IndiaAir$PM_Death_Ambient


China_IndiaAirDeath <- merge(ChinaAir,IndiaAir,by='Year',all = TRUE)
plot(China_IndiaAirDeath$Year,China_IndiaAirDeath$China_outdoor_death,type='l',xlab='year',ylab='Death Caused by outdoor airpollution',col='red',cex=20,main = 'death caused by outdoor air pollution')
lines(China_IndiaAirDeath$Year,China_IndiaAirDeath$SolidFuels_Death_HouseHold.x,col='red3',cex=20,lty='solid')
lines(China_IndiaAirDeath$Year,China_IndiaAirDeath$SolidFuels_Death_HouseHold.y,col='sienna',cex=20,lty='longdash')
lines(China_IndiaAirDeath$Year,China_IndiaAirDeath$India_outdoor_death,col='sienna2',cex=20,lty='longdash')
legend('topright',legend = c('China Air Outdoor pollution Deathrate','China Air Household pollution Deathrate','India Air Household Pollution Death Rate','India Air Outdoor Pollution Death Rate'),col = c('red','red3','sienna','sienna2'),lty=1:3,cex = 0.5)
#I plotted the China plot first and followed by India.I Finally added my legend.
```

## 3. Merge the air pollution data with the gdp data using merge()

This one is pretty strieght forward. I Merged the two dataframes by Country, Code and Year. I excluded the years without pollution data

```{r Q3}
setwd("D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 3/ESP106_week3_lab1")
AirPollu_GDP <- merge(AirPollu,GDP,by=c("Country",'Code','Year'),all.x = TRUE,all.y = FALSE)
```

## 4 Make a plot with two subplots - one showing a scatter plot between log of per-capita GDP (x axis) and indoor air pollution death rate (y axis) and one showing log of per-capita GDP (x axis) and outdoor air pollution (y axis)

For this question, I first partitioned the plot into two parts. I then plotted out the graphs as usual. However, for some reason, the plot that I came up with is different from the example plot shown in class, and I do hope for some advice on why that is the case. Thank you!

```{r Q4}
par(mfrow=c(1,2))
plot(AirPollu_GDP$LogPopGDP,AirPollu_GDP$SolidFuels_Death_HouseHold,xlab = 'log of Population GDP',type='p',pch=20,ylab = "Air Pollution Deathrate(Death per 100,000)",ylim = c(0,250),main = 'indoor air pollution')
plot(AirPollu_GDP$LogPopGDP,AirPollu_GDP$OutDoorPolluDeath,xlab = 'log of Population GDP',type='p',pch=20,ylab = 'Air Pollution Deathrate(Death per 100,000)',ylim = c(0,250),main = 'outdoor air pollution')
```
