---
title: "ESP 106 Final Project"
author: "Yansong Li","Veronica Herrera"
date: "2023-03-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

# Context

```{r World Map}

```

```{r EU/Russian Area Extent}

```

```{r Reliance on Russian Gas}

```

# What are the EU {Countries of Interest( France and Germany)} and Russia's energy import and export prior to the Russian Invasion of Ukraine?
```{r Energy France (Import, Export) by-product}

```

```{r Energy Germany (Import, Export) by product}

```

```{r Energy  Russia (Import, Export) by product}

```

# What are the trends in fossil fuel (coal, oil, and natural gas) energy production, exports, and imports in the European Union since the Russian Invasion of Ukraine? (Feb 2022 to present)

```{r Fossil Fuel Import (3 trend lines - oil vs. gas vs. coal) }

```

```{r Fossil Fuel Export (3 trend lines - oil vs. gas vs. coal)}

```

# What are the trends of renewable energy investments in the COI after the start of the Russian Invasion of Ukraine?

```{r 2022 Fossil Fuel Investment vs Renewable Energy Investment }

```

# What are the trend in renewable energy (Solar, Wind, Hydro) production since the Ukrainian war?

```{r}

```

# Does the Ukrainian war have any impact on nuclear power production in the EU(Focused countries)

```{r}

```

# What is the trend of oil prices and consumption?

```{r}

```

Conclusion

