---
title: "ESP 106 Final Project"
author: "Yansong Li","Veronica Herrera"
date: "2023-03-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Set working directory (be sure to change wording directory on another computer)
knitr::opts_knit$set(root.dir = "D:/UC Davis/Courses and Research/My Courses/ESP 106/Assignment/esp106-Yansong_Li/Project/Data")

```

#Setup

  Before we start, there are a few housekeeping thing to do first. We will need to first download and unzip the data from the drive. the link to the google drive is 
  "https://drive.google.com/drive/folders/1CctyaHwN_xthKQL474wzbHInWXZlHc3b?usp=share_link". 
  
  Besure to download the Editied Data folder. When downloading, please put the zip data filed in a folder called "Data"
  
  We will also need to download some packages need to do the ploting, data organizing, and reading in the data
```{r}

list.files(getwd())
unzip(list.files(getwd()))
list.files(getwd())
list.files("Edited Data")

list.of.packages <- c("ggplot2", "readxl","tidyr","magrittr","dplyr","plotly")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(ggplot2)
library(readxl)
library(tidyr)
library(magrittr)
library(dplyr)
library(plotly)
```
# Introduction
This project will focus on exploring trends in energy investment,prices, production and consumption since the start of the Russia- Ukraine War that started in 2022. Our Analysis will focus on the energy production, import, and export of Russia, and countries of focus within the EU, specifically Germany and France. 

Other countries, such as China, was originally considered, however, there were issues in obtaining time relevant data, which is data from 2021 to the end of 2022. Indeed, when collecting data for this project, finding time relevant data for many topics were challenging, and data are usually more available for Economic Co-operation and Development (OECD) countries than non-OECD ones. 

# Data
We first tried to load all the data thought the google drive, and although some data can be downloaded and read without issue, others are more problematic. Eventrually, we decide its best to use the old fashion zip file method.

We collected those data from IEA or Ember Climate.Org. Most files are organized in a Tidy Data format, but there are some data, such as the "Monthly Oil Statistics","Reliance on Russian Imports" ,and "Gas Trade Flows", that require some reorganizing. One data file called "World Energy Balance Highlights" was particularly difficult to reorganize into a tidy data format, so we use a package call tidyr to help the process. As we looked through the excel filed, we also singled out the countries of interest and put their data in a seperate data sheet in excel to make them easier to work with

## Loading all the data into data frames
### Excel Files

Data:2022 World Energy Investment
```{r Data}
list.files("Edited Data")
World_Energy_Investment <- as.data.frame(read_xlsx("Edited Data/2022 World Energy Investment .xlsx", sheet = "World"))

Europe_Energy_Investment <- as.data.frame(read_xlsx("Edited Data/2022 World Energy Investment .xlsx", sheet = "Europe"))
```

Data:Monthly Oil (Energy Price) Statistics
```{r Data}
#Monthly Energy Price Statistics
Monthly_Oil_Statistics <- as.data.frame(read_xlsx("Edited Data/Monthly Oil Statistics.xlsx", sheet = "Table"))

Germany_Monthly_Oil_Price_Statistics <- as.data.frame(read_xlsx("Edited Data/Monthly Oil Statistics.xlsx", sheet = "Germany Energy Prices"))

France_Monthly_Oil_Price_Statistics <- as.data.frame(read_xlsx("Edited Data/Monthly Oil Statistics.xlsx", sheet = "France Energy Price"))
```

Data:OPEC Reliance on Russia Fuel
```{r Data}
#OPEC Reliance on Russia Fuel
OPEC_Reliance_on_Russian_imports <- as.data.frame(read_xlsx("Edited Data/Reliance on Russian imports.xlsx", sheet = "Reliance_Fuel"))


France_Reliance_on_Russian_imports <- as.data.frame(read_xlsx("Edited Data/Reliance on Russian imports.xlsx", sheet = "French Reliance on Russia"))


Germany_Reliance_on_Russian_imports <- as.data.frame(read_xlsx("Edited Data/Reliance on Russian imports.xlsx", sheet = "Germnay Reliance on Russia"))

```

Data:Gas Trade Flow 
(For this file, only Germany is included in the analysis, because the data shows France doesn't hace a direct gas link with Russia
```{r Data}
Germany_Gas_Trade_Flows <- as.data.frame(read_xls("Edited Data/Gas Trade Flows _17_02_2023.xls", sheet = "To Germany"))
```

Data:World_Energy_Balance
```{r Data}
World_Energy_Balance <- as.data.frame(read_xlsx("Edited Data/World Energy Balances Highlights 2022.xlsx",skip=1, sheet = "TimeSeries_1971-2021"))

France_Energy_Balance <- as.data.frame(read_xlsx("Edited Data/World Energy Balances Highlights 2022.xlsx", sheet = "France"))

Germany_Energy_Balance<- as.data.frame(read_xlsx("Edited Data/World Energy Balances Highlights 2022.xlsx", sheet = "Germany"))
```


###csv file

Data:EU and Countries of Interest Renewables Generation
```{r Data}
EU_Yearly_Renewables_Generation <- as.data.frame(read.csv("Edited Data/EU Yearly Renewables Generation .csv"))

France_Annual_Renewable_Energy_Generation<- as.data.frame(read.csv("Edited Data/France Annual Renewable Energy Generation Data.csv"))

Germany_Annual_Renewable_Energy_Generation_Data <- as.data.frame(read.csv("Edited Data/Germany Annual Renewable Energy Generation Data.csv"))
```

Data:Monthly Energy Statistics
```{r Data}
Monthly_Energy_Statistics <- as.data.frame(read.csv("Edited Data/Monthly Energy Statistics.csv",skip = 8))
```

Data:Monthly Gas Statistics
```{r Data}
Monthly_Gas_statistics <- as.data.frame(read.csv("Edited Data/Monthly Gas statistics.csv"))
```

Data:Monthly OECD Oil Statistics
```{r Data}
Monthly_OECD_oil_statistics <-as.data.frame(read.csv("Edited Data/Monthly OECD oil statistics.csv"))
```

### Tidying up the data
Since some data are not in the tidy format, we are going to use tidyr package to make them into one. (We didn't come accross the package untill later, and we mainly applied it to dataframes that are too combersome to tidy up by hand)

There are 5 data frames that need to be tidied up, including all the Energy Balance Data, Monthly Oil Statistics Data, and OPEC Reliance on Russia Import.

For the reliance dataframes, I also multiplied the original number by 100 to get the percentage
```{r}
# Energy balance Data
colnames(France_Energy_Balance)
France_Energy_Balance <- France_Energy_Balance %>% gather(Year, Energy, 4:54)

colnames(Germany_Energy_Balance)
Germany_Energy_Balance <- Germany_Energy_Balance %>% gather(Year, Energy, 4:54)

colnames(World_Energy_Balance)
World_Energy_Balance <- World_Energy_Balance %>% gather(Year, Energy, 7:57)

# Monthly Oil Statistics
colnames(Monthly_Oil_Statistics)
Monthly_Oil_Statistics <- Monthly_Oil_Statistics %>% gather(Year, Price, 5:101)

#Reliance on Russia Import
colnames(OPEC_Reliance_on_Russian_imports)
OPEC_Reliance_on_Russian_imports <- OPEC_Reliance_on_Russian_imports %>% gather(Year, Reliance, 3:34)


France_Reliance_on_Russian_imports$Reliance <- France_Reliance_on_Russian_imports$Reliance*100
colnames(France_Reliance_on_Russian_imports) <- c("YEAR","COUNTRY","PRODUCT","Reliance %")

Germany_Reliance_on_Russian_imports$Reliance <- Germany_Reliance_on_Russian_imports$Reliance*100
colnames(Germany_Reliance_on_Russian_imports) <- c("YEAR","COUNTRY","PRODUCT","Reliance %")
```


###Deal with NA Values
Some datas contain chatacters that simbolizes NA value in columns and cells. This make later analysis difficult. We will now deal with them by setting these characters in to NA

```{r}
World_Energy_Balance[World_Energy_Balance ==".."] <- NA
Germany_Energy_Balance[Germany_Energy_Balance ==".."] <- NA
France_Energy_Balance[France_Energy_Balance ==".."] <- NA
France_Energy_Balance[France_Energy_Balance =="c"] <- NA
# c stands for confidential

```


# Context

```{r World Map}

```

```{r EU/Russian Area Extent}

```

```{r Reliance on Russian Gas}

```

# What are the EU {Countries of Interest( France and Germany)} and Russia's energy import and export prior to the Russian Invasion of Ukraine?
```{r Energy France (Import, Export) by-product}

```

```{r Energy Germany (Import, Export) by product}

```

```{r Energy  Russia (Import, Export) by product}

```

# What are the trends in fossil fuel (coal, oil, and natural gas) energy production, exports, and imports in the European Union since the Russian Invasion of Ukraine? (Feb 2022 to present)

```{r Fossil Fuel Import (3 trend lines - oil vs. gas vs. coal) }

```

```{r Fossil Fuel Export (3 trend lines - oil vs. gas vs. coal)}

```

# What are the trends of renewable energy investments in Europe? Has the investment trend changed after the start of the Russian Invasion of Ukraine?
For to answer this question, we are going to take a look at the Europe and World Energy Investment data from 2015 to 2022,and see if the onset o the Ukranian War has increased the investment to renewable energy in Europe compared to the global average. It will also be interesting to look at how does the proportion of the "Clean Energy" investment changed over the years.

When looking at the dataset and the IEA Website that it was downloaded from, the IEA didn't quite specify what exactly count as clean energy or renewable energy. Another point that is somewhat unclear is what is the difference between key terms like supply and generation. This unclear definition makes analysis somewhat. 

For the purpose of this project, we will consider renewable energy investment as the combination of Renewables (Supply), Renewables (Generation), Other End-use Renewables. Other type of energy source like "Clean Fuel" was considered, but the term clean is so ambiguous, that makes it unreliable. However, recognizing that aspects like Energy Storage, which is not counted as renewables also plays a role in renewable energy development, we will also still look at the total investment in clean energy in relation to total energy investment.

```{r Europe vs. World Renewable Energy Investment }
#We first try to sum up the renewable energy investment from the three categories
World_Energy_Investment$Renewable <- World_Energy_Investment$`Renewables(Generation)`+World_Energy_Investment$`Renewables (Supply)`+World_Energy_Investment$`Other end-use renewables  (End-use)`

Europe_Energy_Investment$Renewable <- Europe_Energy_Investment$`Renewables(Generation)`+Europe_Energy_Investment$`Renewables (Supply)`+Europe_Energy_Investment$`Other end-use renewables  (End-use)`

#we then make a plot showing the trends of renewable energy investment of Europe and the World

Renewable_Investment <- plot_ly(Europe_Energy_Investment,x= ~Europe_Energy_Investment$Year, y= ~Europe_Energy_Investment$Renewable, type = "scatter",name = 'Europe Renewable Investment', mode = 'lines+markers')
Renewable_Investment <- Renewable_Investment %>% add_trace(y = ~World_Energy_Investment$Renewable, name = 'World Renewable Investment', mode = 'lines+markers') 
Renewable_Investment <- Renewable_Investment %>% layout(title = "Europe and the World Renewable Energy Investment 2015-2022",
                                    yaxis=list(title="$Billions"),
                                    xaxis=list(title="Year"))
Renewable_Investment
htmlwidgets::saveWidget(as_widget(Renewable_Investment), "Renewable_Investment.html")


```

It can be seen in this graph that the world's renewable energy investment start to rise dramatically since 2017, and Europe's investment has remained steady and only increased since 2020. However, this graph might not be very useful if we want to see the change in proportion of the clean energy investment as compared to the total energy investment. Therefore we are going to create another plot

```{r Clean energy and total energy investment in Europe}

Clean_Energy_Investment <- plot_ly(Europe_Energy_Investment, x = Europe_Energy_Investment$Year, y = Europe_Energy_Investment$`Total (billion $2021)`, type = 'bar', name = 'Totale Energy Investment(2021)', marker = list(color = 'blue'))
Clean_Energy_Investment <- Clean_Energy_Investment %>% add_trace(y = Europe_Energy_Investment$`Total Investment of which: Clean energy`, name = 'Clean Energy Investment', marker = list(color = 'green'))
Clean_Energy_Investment <- Clean_Energy_Investment %>% layout(title = "Europe Total and Clean Energy Investment 2015-2022",
         yaxis = list(title = "$Billions"),
         xaxis = list(title = "Year"))
Clean_Energy_Investment
htmlwidgets::saveWidget(as_widget(Clean_Energy_Investment), "Clean_Energy_Investment.html")
```

```{r Clean energy investment as a percentage to total energy investment}
# This plot shows the percentage of clean energy in relation to total energy investment

Europe_Energy_Investment$Clean_Energy_Percentage <- Europe_Energy_Investment$`Total Investment of which: Clean energy`/Europe_Energy_Investment$`Total (billion $2021)`* 100

Clean_Energy_Investment_Percentage <- plot_ly(Europe_Energy_Investment, x = ~Europe_Energy_Investment$Year, y = ~Europe_Energy_Investment$Clean_Energy_Percentage, type = 'bar',
             
             marker = list(color = 'green'
                           ))
Clean_Energy_Investment_Percentage <- Clean_Energy_Investment_Percentage %>% layout(title = "The percentage of total Europe Energy investement as clean energy Investment 2015-2022",
         yaxis = list(title = "% of Totle Energy Investment in Clean Energy"),
         xaxis = list(title = "Year"))
Clean_Energy_Investment_Percentage
htmlwidgets::saveWidget(as_widget(Clean_Energy_Investment_Percentage), "Clean_Energy_Investment_Percentage.html")
```
Finally, it might be interesting to consider what are the differences between fossil fuel investment compared to renewable energy investment through the years. 

To do this will first combine Fossil Fuels without CCUS, fossil Fuels, Oil&gas Upstream and Coal, Oil and Natural gas generation data together, and compare it with the Renewable Energy Investment calculated earlier

```{r Renewable vs. Fossil Fuel Investment}
Europe_Energy_Investment$Fossil <- Europe_Energy_Investment$`Fossil fuels without CCUS (Supply)`+Europe_Energy_Investment$`Fossil fuels(Fuels)`+Europe_Energy_Investment$`Coal (Generation)`+Europe_Energy_Investment$`Oil and natural gas(Generation)`+Europe_Energy_Investment$`Memo: Oil & gas upstream`

Europe_Renewable_VS_Fossil_Energy_Investment <- plot_ly(Europe_Energy_Investment, x = ~Europe_Energy_Investment$Year, y = ~Europe_Energy_Investment$Renewable, type = 'bar',text = Europe_Energy_Investment$Renewable, textposition = 'auto',marker = list(
                           line = list(color = 'white', width = 1.5)), name = 'Renewable Investment')

Europe_Renewable_VS_Fossil_Energy_Investment <- Europe_Renewable_VS_Fossil_Energy_Investment %>% add_trace(y = ~Europe_Energy_Investment$Fossil,text = Europe_Energy_Investment$Fossil, textposition = 'auto',marker = list(
                           line = list(color = 'white', width = 1.5)), name = 'Fossil Fuel Investment')
Europe_Renewable_VS_Fossil_Energy_Investment <- Europe_Renewable_VS_Fossil_Energy_Investment %>% layout(yaxis = list(title = '$Billion'), barmode = 'stack')

Europe_Renewable_VS_Fossil_Energy_Investment <- Europe_Renewable_VS_Fossil_Energy_Investment %>% layout(title = "Europe Renewable Vs. Fossil Fuel Investment",
         xaxis = list(title = "Year"),
         yaxis = list(title = "$Billion"))
htmlwidgets::saveWidget(as_widget(Europe_Renewable_VS_Fossil_Energy_Investment), "Europe_Renewable_VS_Fossil_Energy_Investment.html")
Europe_Renewable_VS_Fossil_Energy_Investment
```

Findings:
As we see from the graph presented above. Europe's renewable energy investment is at a steady state, without experiencing the drastic increase in renewable investment that over took the world since 2017. Although the money diverted to  clean energy saw a marked increase in 2021 and 2022, the total investment in the energy system also increase. Which means that the proportion of clean energy investment within the total energy investment didn't change that much. Finally, we can see that since 2015, renewable energy sector investment has increase relative to Fossil Fuels. From 2015 to 2019, Fossil Fuel investment always exceeded renewable energy investment, but since 2020, the opposite happened. However, it is important to note that despite renewable energy investment does surpass those of fossil fuels, ratio between the two doesn't seem close to the 68% to 80% of clean energy investment proportion to the total energy investment, which begs the question of what other energy investments are being counted as clean energy investment apart from those of renewable energy?

As for the impact of Ukranian War, the graph doesn't seem to show any drastic increase in both renewable or the broader defined clean energy investment in Europe since the start of the war. Perhaps this is because the war only happened recently, and its full impact on energy investment is still unfolding.


# What are the trend in renewable energy (Solar, Wind, Hydro) production since the Ukrainian war?

```{r Europe Renewable Energy Production} 

```


# Does the Ukrainian war have any impact on nuclear power production in the EU(Focused countries)

```{r}
```

# What is the trend of oil prices and consumption?

```{r}

```

Conclusion

