---
title: "Lab 6"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

path <- "...."

knitr::opts_knit$set(root.dir = path)
setwd(path)
```

## Before you start

Before you start, install the current version of the "terra" package *like this:*

```{r}
install.packages('terra', repos='https://rspatial.r-universe.dev')
```

(unless it asks you to install from source code, in that case use the normal install)

Read a [short introduction](https://rspatial.org/spatial/index.html) to spatial data in R.  

There are a lot of questions. At first, do not spend too much time on making the perfect maps (you can come back to that later)

## Introduction

In this lab we will use some GPS tracking data for Blue Whales in the eastern North Pacific, downloaded from Movebank <https://www.movebank.org> which has lots of wildlife tracking data. We will make some maps and overlay polygons of Marine Protected Areas off the California coast to understand how whales move into and out of MPAs.

## Whale data

1. Use `terra::vect` to read in the datapoints for the Blue Whale migration data ("points.shp"). 

```{r}
library(terra)
setwd("D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 6/Blue Whale Migration Data")
bluew <- vect("points.shp")
bluew
```

This is a set of point location for GPS tagged blue whales the eastern north Pacific. Identifiers for individual whales are given in the ind_ident column. 

2) How many observations are there in this dataset? And how many individual whales are tracked? 

```{r}
Bluew <- as.data.frame(bluew)
nrow(Bluew)
nindividual <- as.factor(Bluew[,'ind_ident'])
nindividual
nlevels(nindividual)
#There are 15545 different observations recorded, and 141 blue whales studied 
```

## Mapping whales

3. Get the global country boundaries with the geodata package 

```{r}
install.packages('geodata')
wrld <- geodata::world(path="D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 6/Blue Whale Migration Data")
```


4. Make a plot showing the whale observations, using a different color for each whale (many colors will be alike). Add the west coast of North America for reference. Set the limits of the plot to the extent of the whale migration data. 

Refine the below. Plot 1 on canvas shows you what you can aim for.

```{r}
# Step 1. plot the countries using the extent of the whale data to only show the area of interest. 
library(terra)
# the area of interest
bluew <- vect("points.shp")
aoi <- ext(bluew) + 2
plot(wrld, ext=aoi)

#Step 2: plot the whale points. You can use either the "points" method or "plot( , add=TRUE)". The latter is easier for coloring, but you need to tell it not to add a legend.

plot(bluew,col=rainbow(nlevels(nindividual)), add=TRUE)

```

5) make new SpatVectors for each of the following two individuals: "2008CA-Bmu-00825" and "2005CA-Bmu-10821"

```{r}
bw00825 <- bluew[bluew$ind_ident=="2008CA-Bmu-00825"]
BW00835 <- as.data.frame(bw00825)
bw10821 <- bluew[bluew$ind_ident=="2005CA-Bmu-10821"]
BW10821 <- as.data.frame(bw00825)
```

6) Combine these two into one new SpatVector 

```{r}
TwoBW <- rbind(bw00825,bw10821)
BWs <- as.data.frame(TwoBW)
#I am not sure if this is the way to do it
```


7) Make a map showing the tracks of these two using different colors. See plot2 on Canvas for an example.

```{r}
aoi <- ext(bluew) + 2

plot(wrld, ext=aoi)

plot(TwoBW,col=rainbow(2), add=TRUE)
```

## Protected areas

8) The folder MPAs has a shapefile in it with polygons for Marine Protected Areas within the United States. This file was extracted from a file with all protected areas in the US [source](https://www.protectedplanet.net). Read the MPA shapefile. 

```{r}
library(terra)
setwd('D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 6/MPAs')
list.files('D:/UC Davis/Courses and Research/My Courses/ESP 106/Lab/Lab 6/MPAs')
mpa <- vect("mpas.shp")
mpa
plot(mpa)
```

9) Now we will focus on just the US west coast. You can use the following longitude and latitude extent: xmin=-128, xmax=-115, ymin=32, ymax=40. Get just the MPAs on the US west coast by cropping the MPA dataset to this extent.

```{r}
#Hint: use the crop() function
e <- ext(-128,-115,32,40)
mpa <- crop(mpa,e)
mpa
plot(mpa)
```

10) Find the fraction of the blue whale observations that are in one of the west coast MPAs. You can use `is.related`

```{r}
t <- is.related(bluew,mpa,"within")
table(t)["TRUE"]
#2496 of the 15545 observations(16%) are with in one of the west coast MPAs.
```

11) Find what fraction of blue whales in the dataset spend at least some time in one of these west coast MPAs. 

```{r}
Bluew$Spend_TimeWC <- t
WCSpend_Time <- Bluew[Bluew$Spend_TimeWC=="TRUE",]
nlevels(as.factor(WCSpend_Time$ind_ident))/nlevels(as.factor(Bluew$ind_ident))*100
#there are 119 blue whale that spent time in the wet coast, and there are 141 blue whale in total. This means that 84.397% of the all blue whales had spent sometime in west coast.
```


12. Create a raster with counts of blue whale observations along the California coast. Use the spatial extent we used for the mpa data (question 8) and use a spatial resolution of 1/6th of a degree. First create an empty SpatRaster. Then use `rasterize`

```{r}
r <- rast(e, res=1/6)
r <- rasterize(bluew,r,fun="count")
r
plot(r)
```

13. Make a map to show the counts and the MPAs and land areas. Plot 3 on Canvas shows you the plot you are aiming for. 

```{r}
plot(r,add=TRUE)
plot(wrld, ext=e,col="grey",add=TRUE)
plot(mpa,ext=e,border="blue", add=TRUE)

```

14) Using the country boundaries, compute for the raster with the number of observations, the distance to the coast

```{r}
#I'm not sure how to do this one. I started researching on google about how to calculate distance but everything I found was on how to calculate distance between two points, bot between a raster and a vector. Maybe its about changing the raster data into a vector? After not being able to find a solution, I start to work with Veronica, but she also couldn;t figur out a way to do this question.
```

15) For each blue whale observation, find the distance to the coast based on this raster (use "extract"). Make a histogram of the distribution. 

```{r}
# Same as the above.
```

